---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---

by Tara Lutman Agacayak

github repo with RMarkdown source code: https://github.com/tagacayak/RepData_PeerAssessment1

#Loading and preprocessing the data

```{r step1}
## Unzip & read file

activityData <- read.csv("activity.csv")
```

#What is mean total number of steps taken per day?

1. Make a histogram of the total steps taken per day.

```{r step2.1}
stepsPerDay <- aggregate(steps ~ date, data=activityData, FUN=sum)
hist(stepsPerDay$steps, main = "Histogram of Total Steps Per Day", xlab = "steps per day")
```

2. Calculate and report the **mean** and **median** total number of steps taken per day

```{r step2.2}
dailyMean <- mean(stepsPerDay$steps)
dailyMedian <- median(stepsPerDay$steps)
```

The mean total number of steps taken per day is `r dailyMean`

The median total number of steps taken per day is `r dailyMedian`

#What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r step3.1}
avgStepsPerInterval <- aggregate(steps ~ interval, data=activityData, mean)
plot(avgStepsPerInterval, type="l")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r step3.2}
maxInterval <- which.max(avgStepsPerInterval$interval)
```

The interval which contains the maximum number of steps is interval `r maxInterval`.

#Imputing missing values

```{r step4}
plot(cars)
```

#Are there differences in activity patterns between weekdays and weekends?

```{r step5}
plot(cars)
```